---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# feedeffir

<!-- badges: start -->
<!-- badges: end -->

The goal of feedeffir is to process feed efficiency files and calculate total dry matter intake, metabolic body weight, delta body weight, and milk energy.

## Installation

You can install the development version of feedeffir from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("GMBog/feedeffir")
```

feedeffir handles two main types of data:

1. **Raw Data**: 
   - **Description**: This includes unprocessed datasets directly from the farm. Examples are raw milk weights from DC and AFI system, feed intake files from RIC2 Discover system, and milk composition files.
   - **Usage**: Functions that operate on raw data are designed for pre-processing and compilation.

2. **Compiled Feed Efficiency Data**:
   - **Description**: This includes compiled datasets from raw data and include metrics such as milk weights, milk composition, and feed intakes.
   - **Usage**: Functions targeting compiled files focus on analyzing feed efficiency traits. They require data that has already been processed from raw inputs.

### How to Use

To work with **raw data**, use functions to preprocess and create the initial data sets:

* `process_VRfiles()`
* `process_DC_milkw()`
* `process_AFI_milkw()`

To compile data, use the compilers:

* `compile_bw_files()`
* `compile_milkw_files()`
* `compile_milkcomp_files()`

To analyze **compiled feed efficiency files**, use functions which expect compiled data:

* `process_bw()`
* `process_intakes()`
* `calculate_milke()` 

For detailed examples of how to use these functions, refer to the [Examples](#examples) section.


## Examples

This is a basic example which shows you how to solve a common problem:

```{r example}
library(feedeffir)

# Process VR files with intakes
file_path <- system.file("extdata", "VR240724.DAT", package = "feedeffir")

data <- process_VRfiles(
  exp = "Study1",
  VRfile = file_path,
  bins = seq(1, 32),
  output_dir = tempdir()
)

head(data)

# The next step is to compile the processed VR files
# compile_VRfiles(dir = "~/Downloads/files/",
#                compfile = "UW_Study1_CompiledIntakes.xlsx",
#                data = data)



# Process milk weights from Dairy Comp
exp <- "Study1"
file_path <- system.file("extdata", "MilkWeight_DCfile.xls", package = "feedeffir")
output_dir <- tempdir()
data <- process_DC_milkw(exp, file_path, output_dir)

head(data)
```

